CF-SuGAR Power Calculations

### 2025-03-19 Christine Chan/CF-SuGAR/power_calculations.qmd
base models for Deb

  Age_YrEnd = "Age", bmi_perc = "BMI %ile"

@@

@@

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Need to use a dataframe not a tibble
annual <- as.data.frame(annual)
annual$tstart <- as.numeric(annual$tstart) - as.numeric(og)
annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
# Formula for base model
xvars <- c(
  "Age_YrEnd", "Gender", "bmi", "MutClass", "tube_feeding", "liver_disease", "ppFEV1_cat",
  "NumPulmExacerbation"
)
y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

## Base model plus HbA1c

```{r}
# Formula for base model
cox_res <- coxph(update(f, . ~ . + A_hgba1c), data = annual, cluster = eDWID)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

# 1 hour OGTT glucose in CFRD

## 1 hour glucose alone

@@

ogtt <- read.csv("./1 hour OGTT in CFRD/Data_raw/ogtt.csv", na.strings = "")
pft <- read.csv("./1 hour OGTT in CFRD/Data_raw/pfts.csv", na.strings = "")
a1c <- read.csv("./1 hour OGTT in CFRD/Data_raw/a1c.csv", na.strings = "")
demo <- read.csv("./1 hour OGTT in CFRD/Data_raw/demographics.csv", na.strings = "")
dob <- read.csv("./1 hour OGTT in CFRD/Data_raw/dob.csv", na.strings = "")

@@

# write file of people with missing genotype risk
# write.csv(alldata[alldata$genoRisk=="",],"./Christine Chan/1 hour OGTT in CFRD/Data_raw/missing_genotype_risk.csv")

@@

# checking ppts that Collin said had CFRD
col_check <- alldata[alldata$MRN %in% c(
  1006479, 1053930, 1078408, 1136992, 1289450, 1339023, 1413821, 1695512, 664400,
  686476, 707831, 712676, 739231, 739460, 787372, 800007, 848440, 860181, 875128, 897478, 905284, 908966,
  918124, 932695, 949042, 952250, 952252, 825118
), ]
col_check <- col_check[order(col_check$MRN, col_check$Date), ]
# write.csv(col_check,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd.csv",na="",row.names = F)

# last 2 discrepant cases after fixing date issue
col_check2 <- alldata[alldata$MRN %in% c(686476, 825118), ]
col_check2 <- col_check2[order(col_check2$MRN, col_check2$Date), ]
# write.csv(col_check2,"./Christine Chan/1 hour OGTT in CFRD/Data_raw/check_cfrd2.csv",na="",row.names = F)

@@

##############
# COX MODELS #
##############
baseline_mult$Race <- factor(baseline_mult$Race, levels = c("Caucasian", "Black", "Other"))
# CPH model for baseline 1 hour glucose
f <- Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + FEV1.per.pred + Baseline_1hr_glucose
cox1hr <- coxph(f, data = baseline_mult)
# Forest plot and coefficients table
cox1hr %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox1hr,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

## 1 hour glucose and HbA1c

```{r}
cox1hra1c <- coxph(update(f, . ~ . + A1C), data = baseline_mult)
# Forest plot and coefficients table
cox1hra1c %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox1hra1c,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

- There are no participants over 31 in Christine's 1 hour OGTT analysis data, so I used age as a continuous variable.
- Because these models include age and sex, I also used raw BMI value as a continuous variable.
- Race was removed from the models because the group sizes are so small in the one hour dataset (only 3 Black participants).
- Pancreatic status was not available in the one hour glucose dataset, so was removed from the models.

# Questions for Deb

- Part of the reason that we split the data into 3 eras was because the PH assumption was violated in the full model. How do you want to handle that for the power calculations?

>>>>

  Age_YrEnd = "Age", bmi_perc = "BMI %ile",
  A_hgba1c = "HbA1c"

@@

# Limit to after 2010
annual <- annual %>% filter(ReviewYear >= 2010)

@@

```{r}
# Origin for survival analyses
og <- ymd("2003-01-01")
# Need to use a dataframe not a tibble
annual <- as.data.frame(annual)
annual$tstart <- as.numeric(annual$tstart) - as.numeric(og)
annual$tstop <- as.numeric(annual$tstop) - as.numeric(og)
# Formula for base model
xvars <- c(
  "age_group", "Gender", "Race", "Hispanicrace", "bmi", "MutClass",
  "A_IsOnEnzymes", "tube_feeding", "liver_disease", "A_FEV1_pct_predicted",
  "NumPulmExacerbation", "A_corticosteroids1", "A_hgba1c"
)
y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
cox.zph(cox_res)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

## Base model with variables available in Lorenz et al.

```{r}
# Formula for base model
xvars <- c(
  "age_group", "Gender", "bmi", "MutClass", "A_FEV1_pct_predicted", "A_hgba1c"
)
y <- "Surv(time = tstart, time2 = tstop, event = as.numeric(CFRD) == 2)"
f <- as.formula(paste0(y, "~", paste0(xvars, collapse = "+")))
cox_res <- coxph(f, data = annual, cluster = eDWID)
cox.zph(cox_res)
# Forest plot and coefficients table
cox_res %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox_res,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

# Lorenz et al.

@@

ogtt <- read.csv("./1 hour OGTT in CFRD/Data_raw/OGTT.csv")
pft <- read.csv("./1 hour OGTT in CFRD/Data_raw/PFTs.csv")
a1c <- read.csv("./1 hour OGTT in CFRD/Data_raw/A1c.csv")
demo <- read.csv("./1 hour OGTT in CFRD/Data_raw/demographics.csv")
dob <- read.csv("./1 hour OGTT in CFRD/Data_raw/dob.csv")

@@

# write file of people with missing genotype risk
# write.csv(alldata[alldata$genoRisk=="",],"./1 hour OGTT in CFRD/Data_raw/missing_genotype_risk.csv")

@@

# checking ppts that Collin said had CFRD
col_check <- alldata[alldata$MRN %in% c(
  1006479, 1053930, 1078408, 1136992, 1289450, 1339023, 1413821, 1695512, 664400,
  686476, 707831, 712676, 739231, 739460, 787372, 800007, 848440, 860181, 875128, 897478, 905284, 908966,
  918124, 932695, 949042, 952250, 952252, 825118
), ]
col_check <- col_check[order(col_check$MRN, col_check$Date), ]
# write.csv(col_check,"./1 hour OGTT in CFRD/Data_raw/check_cfrd.csv",na="",row.names = F)

# last 2 discrepant cases after fixing date issue
col_check2 <- alldata[alldata$MRN %in% c(686476, 825118), ]
col_check2 <- col_check2[order(col_check2$MRN, col_check2$Date), ]
# write.csv(col_check2,"./1 hour OGTT in CFRD/Data_raw/check_cfrd2.csv",na="",row.names = F)

@@

##############
# COX MODELS #
##############
```

```{r}
# CPH model for baseline 1 hour glucose
cox1hr <- coxph(
  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + class1 + FEV1.per.pred +
    A1C,
  data = baseline_mult
)
# Forest plot and coefficients table
cox1hr %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox1hr,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

## Model with one hour OGTT

```{r}
# CPH model for baseline 1 hour glucose
cox1hr <- coxph(
  Surv(Follow_up_time_yrs, CFRD) ~ age + Sex + BMI + class1 + FEV1.per.pred +
    A1C + Baseline_1hr_glucose,
  data = baseline_mult
)
# Forest plot and coefficients table
cox1hr %>%
  tbl_regression(exponentiate = T)
ggcoef_table(cox1hr,
  show_p_values = F, add_reference_rows = T, exponentiate = T,
  signif_stars = F, colour = F
)
```

@@

- I removed race from the models because the Lorenz et al. paper data only includes 8 non-white participants, so the coefficients exploded.
